---
title: "ReadMe"
author: "Dean Watson"
date: "3/4/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##  Peer-graded Assignment: Getting and Cleaning Data Course Project
##  Due March 11,2018
##  Submitted by Dean Watson
##  in Credit for Data Science Certification
##
The purpose of this project is to demonstrate your ability to collect, work with, 
and clean a data set. The goal is to prepare tidy data that can be used for later 
analysis. You will be graded by your peers on a series of yes/no questions related 
to the project. You will be required to submit: 1) a tidy data set as described below, 
2) a link to a Github repository with your script for performing the analysis, and 
3) a code book that describes the variables, the data, and any transformations or 
work that you performed to clean up the data called CodeBook.md. You should also 
include a README.md in the repo with your scripts. This repo explains how all of the 
scripts work and how they are connected.

    Review Criteria
    1. The submitted data set is tidy.
    2. The Github repo contains the required scripts.
    3. GitHub contains a code book that modifies and updates the available codebooks with the data to indicate all the variables and summaries calculated, along with units, and any other relevant information.
    4. The README that explains the analysis files is clear and understandable.
    5. The work submitted for this project is the work of the student who submitted it.

 The following scripd accomplishes the following steps:
    1. Merges the training and the test sets to create one data set.
    2. Extracts only the measurements on the mean and standard deviation for each measurement.
    3. Uses descriptive activity names to name the activities in the data set
    4. Appropriately labels the data set with descriptive variable names.
    5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
if(!file.exists("./data")){dir.create("./data")}
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
download.file(fileUrl,destfile="./data/Dataset.zip",method="curl")
unzip(zipfile="./data/Dataset.zip",exdir="./data")
path_rf <- file.path("./data" , "UCI HAR Dataset")
files<-list.files(path_rf, recursive=TRUE)
```
Values of Varible Activity consist of data from “Y_train.txt” and “Y_test.txt”
   1. values of Varible Subject consist of data from “subject_train.txt” and subject_test.txt"
   2. Values of Varibles Features consist of data from “X_train.txt” and “X_test.txt”
   3. Names of Varibles Features come from “features.txt”
   4. levels of Varible Activity come from “activity_labels.txt”
  We will use Activity, Subject and Features as part of descriptive variable names for data
```{r}
# Read Activity Data 
activityTestData  <- read.table(file.path(path_rf, "test" , "Y_test.txt" ),header = FALSE)
activityTrainData <- read.table(file.path(path_rf, "train", "Y_train.txt"),header = FALSE)

# Read Subject Data 
subjectTrainData <- read.table(file.path(path_rf, "train", "subject_train.txt"),header = FALSE)
subjectTestData  <- read.table(file.path(path_rf, "test" , "subject_test.txt"),header = FALSE)

# Read Subject Data 
featuresTestData  <- read.table(file.path(path_rf, "test" , "X_test.txt" ),header = FALSE)
featuresTrainData <- read.table(file.path(path_rf, "train", "X_train.txt"),header = FALSE)

# Concatenate the test and train tables by rows
subjectData <- rbind(subjectTrainData, subjectTestData)
activityData<- rbind(activityTrainData, activityTestData)
featuresData<- rbind(featuresTrainData, featuresTestData)

## set names to variables
names(subjectData)<-c("subject")
names(activityData)<- c("activity")
featuresNamesData <- read.table(file.path(path_rf, "features.txt"),head=FALSE)
names(featuresData)<- featuresNamesData$V2

## Merge columns to get the data frame Data for all data
combineData <- cbind(subjectData, activityData)
Data <- cbind(featuresData, combineData)

## Subset Name of Features by measurements on the mean and standard deviation( used grep)
subFeaturesNamesData<-featuresNamesData$V2[grep("mean\\(\\)|std\\(\\)", featuresNamesData$V2)]

## Subset the data frame Data by seleted names of Features
selectedNames<-c(as.character(subFeaturesNamesData), "subject", "activity" )
Data<-subset(Data,select=selectedNames)

## Read activity names
activityLabels <- read.table(file.path(path_rf, "activity_labels.txt"),header = FALSE)

Data$activity<-factor(Data$activity);
Data$activity<- factor(Data$activity,labels=as.character(activityLabels$V2))

```
  Appropriately labels the data set with descriptive variable names
 prefix t  is replaced by  time
 Acc is replaced by Accelerometer
 Gyro is replaced by Gyroscope
 prefix f is replaced by frequency
 Mag is replaced by Magnitude
 BodyBody is replaced by Body 
```{r}
names(Data)<-gsub("^t", "time", names(Data))
names(Data)<-gsub("^f", "frequency", names(Data))
names(Data)<-gsub("Acc", "Accelerometer", names(Data))
names(Data)<-gsub("Gyro", "Gyroscope", names(Data))
names(Data)<-gsub("Mag", "Magnitude", names(Data))
names(Data)<-gsub("BodyBody", "Body", names(Data))

```
  Creates a second,independent tidy data set and ouput it
In this part,a second, independent tidy data set will be created with the average 
of each variable for each activity and each subject  based on the data set in step 4. 
```{r}
library(plyr);
Data2<-aggregate(. ~subject + activity, Data, mean)
Data2<-Data2[order(Data2$subject,Data2$activity),]
write.table(Data2, file = "tidydata.txt",row.name=FALSE)

```




